Multiple Dice Working as One: CAD Flows and Routing Architectures for Silicon Interposer FPGAs

### 2015-03-04 doc/journal/fpgapaperfinal/FPGAPaper/partitioning.tex
Reworded issues to issue, re: metis vs hmetis
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4588 8e3573b8-cf2c-4f14-ef6d-137439e28b8b
Commented on lack of optimality of random partition assignment
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4589 8e3573b8-cf2c-4f14-ef6d-137439e28b8b
Commented on origin of edge weight schemes
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4590 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

\subsection{Partitioner Stage}\label{sec:partitioner_stage}
As shown in Figure \ref{fig:partitioning_cad_flow}, we explore the effect of circuit partitioning at two different stages of the CAD flow. The first possibility is to partition the netlist of primitive elements (LUTs, FFs, RAM slices, etc.) before packing into the larger function blocks such as logic blocks and (32 kb) RAM blocks. The partitioning constraints must be respected by the packer; it must not pack two primitives assigned to different partitions (dice) into the same function block. We updated the VPR clustering algorithm to respect this new constraint. The second possible way to incorporate partitioning is after packing, where partitioning is applied to the function block (cluster-level) netlist. Each partition is then assigned (randomly) to a separate FPGA die. The partitioning constraints must then be respected by the placer; it is free to move function blocks around within a die, but not across interposer cutlines. These two partitioning flow options can be combined -- if partitioning is applied before packing, one can also pass the partition information to the placement stage. The post-packing placement constraints can be derived from the primitive block partitioning constraints, with the valid placement region for a given function block being equal to the intersection of the valid placement regions of the primitives packed into that function (clustered) block. There are, therefore, four possible partitioning flows:

@@

\subsection{Partitioning Tool}
Two well-known freely available graph partitioning tools are Metis~\cite{karypis1998multilevelmetis} and hMetis~\cite{karypis1999multilevelhmetis}. Metis operates on graphs, while hMetis operates on hypergraphs -- a hypergraph is a generalization of a graph, where each edge can connect any number of vertices. It is most natural to represent a circuit netlist as a hypergraph, where the vertices are netlist pins (or blocks) and the hyperedges are the nets joining the pins. Owing to the natural hypergraph representation, hMetis would be the most appropriate choice of partitioner. However, there are other issues to consider for our use case.

@@

We explored several ways to generate the per-net subgraphs. The two parameters in this operation are the graph topology and the edge weight scheme. We considered two graph topologies, clique and star, as illustrated in Figure \ref{fig:star_clique}. We assigned the same edge weight to every edge of the generated subgraph. Edge weights were computed based on the number of vertices in the originating hyperedge, $n$, and we considered edge weights equal to $1$, $1/n$, and $1/n^2$. We use \emph{hyperedge model} to refer to the combination of graph topology and edge weight scheme.

>>>>

\subsection{Partitioner Stage}\label{sec:partitioner_stage}
As shown in Figure \ref{fig:partitioning_cad_flow}, we explore the effect of circuit partitioning at two different stages of the CAD flow. The first possibility is to partition the netlist of primitive elements (LUTs, FFs, RAM slices, etc.) before packing into the larger function blocks such as logic blocks and (32 kb) RAM blocks. The partitioning constraints must be respected by the packer; it must not pack two primitives assigned to different partitions (dice) into the same function block. We updated the VPR clustering algorithm to respect this new constraint. The second possible way to incorporate partitioning is after packing, where partitioning is applied to the function block (cluster-level) netlist. Each partition is then assigned (randomly) to a separate FPGA die. Our use of random assignment is not optimal, and an interesting area for future work is to explore keeping strongly connected partitions in adjacent dice. The partitioning constraints must then be respected by the placer; it is free to move function blocks around within a die, but not across interposer cutlines. These two partitioning flow options can be combined -- if partitioning is applied before packing, one can also pass the partition information to the placement stage. The post-packing placement constraints can be derived from the primitive block partitioning constraints, with the valid placement region for a given function block being equal to the intersection of the valid placement regions of the primitives packed into that function (clustered) block. There are, therefore, four possible partitioning flows:

@@

\subsection{Partitioning Tool}
Two well-known freely available graph partitioning tools are Metis~\cite{karypis1998multilevelmetis} and hMetis~\cite{karypis1999multilevelhmetis}. Metis operates on graphs, while hMetis operates on hypergraphs -- a hypergraph is a generalization of a graph, where each edge can connect any number of vertices. It is most natural to represent a circuit netlist as a hypergraph, where the vertices are netlist pins (or blocks) and the hyperedges are the nets joining the pins. Owing to the natural hypergraph representation, hMetis would be the most appropriate choice of partitioner. However, another important issue is the resource balancing ability of each tool.

@@

We explored several ways to generate the per-net subgraphs. The two parameters in this operation are the graph topology and the edge weight scheme. We considered two graph topologies, clique and star, as illustrated in Figure \ref{fig:star_clique}. We assigned the same edge weight to every edge of the generated subgraph. Edge weights were computed based on the number of vertices in the originating hyperedge, $n$, and we considered edge weights equal to $1$, $1/n$, and $1/n^2$.  Our intuition is that assigning lower edge weight to high-fanout nets may be beneficial, and we selected the edge weights accordingly. We use \emph{hyperedge model} to refer to the combination of graph topology and edge weight scheme.
